loading the data from the url given.

```{r, echo = TRUE}
temp <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
data <- read.csv(unz(temp, "activity.csv"))
unlink(temp)
```

Processing the activity data

```{r, echo=TRUE}
day<- weekdays(as.Date(data$date))
date_time<- as.POSIXct(data$date, format="%Y-%m-%d")
```

pulling data without nas

```{r, echo= TRUE}
clean_data <- data[!is.na(data$steps),]
```

Histogram of total number of steps taken each day.

Calculating the total number of steps taken each day.

```{r, echo= TRUE}
sum1 <- aggregate(data$steps~data$date,FUN=sum,)
colnames(sum1) <- c("date","steps")
hist(sum1$steps, col= "red",xlab= "steps", main= "Total steps taken in a day")
```

mean and median of the number of steps taken each day.

```{r, echo= TRUE}
mean1 <- mean(sum1$steps)
median1 <-median(sum1$steps)
````
the mean of the steps is 10766.19
the median of the steps is 10765


time series plot of the average number of steps taken.
creating the intervals for average number of steps taken.

```{r, echo= TRUE}
library(plyr)
library(ggplot2)
interval_data <- ddply(clean_data, .(interval), summarize, avg = mean(steps))
g <- ggplot(interval_data, aes(x=interval, y= avg), xlab= "interval", ylab= "
average number of steps")
g+geom_line()+ggtitle("Average number of steps per interval")
```

Maximum number of steps per interval.
``` {r, echo= TRUE}
maximum <- max(interval_data$avg)
interval_data[interval_data$avg==maximum,1]
```
the interval with maximum  number of steps is 835  

Code to describe and show a strategy for imputing missing data.


```{r, echo=TRUE}
steps_by_interval <- aggregate(steps ~ interval, data, mean)
imputed_data <- transform(data, steps = ifelse(is.na(data$steps), steps_by_interval$steps[match(data$interval, steps_by_interval$interval)], data$steps))
```

Histogram to show the difference
```{r, echo= TRUE}
steps_by_day_i <- aggregate(steps ~ date, imputed_data, sum)
hist(steps_by_day_i$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")
hist(sum1$steps, main = paste("Total Steps Each Day"), col="red", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("blue", "red"), lwd=10)
```

Calculating mean and median
```{r, echo= TRUE}
mean2 <- mean(steps_by_day_i$steps)
median2 <- median(steps_by_day_i$steps)
mean2 - mean1
median2 - median1
sum(steps_by_day_i$steps) - sum(sum1$steps)
```
the mean of the steps is 10766.19  
the median of the steps is 10766.19  

Check the difference   
the difference between the means is 0  
the difference between the medians is 1.188679  
the total difference between the steps is 86129.51  

Activity patterns in weekdays and weekends

```{r, echo= TRUE}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
imputed_data$dow = as.factor(ifelse(is.element(weekdays(as.Date(imputed_data$date)),weekdays), "Weekday", "Weekend"))
steps_by_interval_i <- aggregate(steps ~ interval + dow, imputed_data, mean)
library(lattice)

xyplot(steps_by_interval_i$steps ~ steps_by_interval_i$interval|steps_by_interval_i$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")

```


